### 2019-12-18 defesa/tex/chapter/e3p.tex
b1

The problem of finding every center and angle of rotation of a fixed shape ellipse which makes it have three points on its border is presented in this section. Even though its simple statement--it is short and uses only basic mathematical concepts--we were not able to find any work on it, or even on related problems. 
As a result, starting from scratch, we ended up trying a handful of approaches with most of them failing on the way. We try to give a review of some of those, going through the issues with the failing ones. At the end we propose a $\bigO(1)$ algorithm that finds every solution of the problem, and also run some experiments to analyze efficiency in terms of numerical precision and CPU time.

We refer to this problem as \sigla{E3PNT}{Ellipse by Three points}, and an instance of it is given by three points $u, v, w \in \R^2$ and $E$, the coverage area of an ellipse with shape parameters $(a, b) \in \R^2_{>0}$, with $a > b$. A solution of E3PNT is a pair $(q, \theta) \in \R^2\times[0, \pi)$, such that $\{u, v, w\} \subset \partial E(q, \theta)$. Because of its application on \autoref{chapter:ellipses_n}, every solution of E3PNT must be encountered. This is a vital piece of information which makes this problem more challenging.

@@

@@

It is possible to make another reduction to obtain a degree-$6$ polynomial, this can lower the size of the companion matrix utilized on obtaining the roots of $h$ from $12^2=144$ to $6^2=36$.
As it has been mentioned before, an ellipse is symmetric with respect to its own major or minor axis. 
In other words, rotating an ellipse by $\theta \in [0, \pi)$ is the same as rotating the same ellipse by $\pi + \theta$.
On the other hand, on \autoref{chapter:definitions}, it was stated that the angle of the complex numbers $z$ and $-z$ has the same symmetry with each other as the aforementioned symmetry of ellipses:
\begin{equation*}
angle(-z) = \pi + angle(z).
\end{equation*}
From that, as $g(e^{i\theta})=\xi(\theta)$ for every $\theta \in [0, 2\pi)$, we conclude that $g(-z)=g(z)$. This implies that $h$ is, in fact, an even polynomial, or that $h(-z) = h(z)$ is true for every $z\in\mathbb{C}$:
\begin{align}
h(-z) = (-z)^6g(-z) = z^6g(z).
\end{align}
Therefore, all the odd degree coefficients of $h$ are $0$ and we can define the $6$-degree polynomial $f : \mathbb{C} \mapsto \mathbb{C}$ with the substitution $y=z^2$:
\begin{equation}
f(y) = \sum_{k=0}^{6} c_{2k} y^k.
\end{equation}
Then from every root $\hat{y}$ of $f$, two roots of $h$ can be obtained: $\sqrt{\hat{y}}$ and $-\sqrt{\hat{y}}$. As the angle of $-\sqrt{\hat{y}}$ is not between $[0, pi)$ we can ignore it. Note that the the square root of $\hat{y}$ does not need to be calculated, as only the angles are needed and they can be obtained by:
\begin{equation}
angle(\sqrt{z}) = angle(z)/2.
\end{equation}
Finally, using the QR algorithm mentioned on \autoref{chapter:definitions} a $\bigO(n^3)$ algorithm, with $n=6$, can be constructed for E3P.

>>>>

The problem of finding every center and angle of rotation of a fixed shape ellipse which makes it have three points on its border is presented in this section. Even though its simple statement--it is short and uses only basic mathematical concepts--we were not able to find any work on it, or even on related problems. 
As a result, starting from scratch, we ended up trying a handful of approaches with most of them failing on the way. We try to give a review of some of those, going through the issues with the failing ones. At the end we propose an algorithm that takes around $6^3$ operations to compute every solution of the problem. Also, we run some experiments to analyze its efficiency in terms of numerical precision and CPU time.

We refer to this problem as \sigla{E3PNT}{Ellipse by Three points}, and an instance of it is given by three points $u, v, w \in \R^2$ and $E$, the coverage area of an ellipse with shape parameters $(a, b) \in \R^2_{>0}$, with $a > b$. A solution of E3PNT is a pair $(q, \theta) \in \R^2\times[0, \pi)$, such that $\{u, v, w\} \subset \partial E(q, \theta)$. As a last remark, because of its application on \autoref{chapter:ellipses_n}, we need to encounter every solution of E3PNT, this requirement, of course, makes the problem more challenging.

@@

Initially, E3PNT is a problem with $3$ unknown variables: $q_x, q_y, $ and $\theta$. In this section, we transform E3PNT into the problem of finding the roots of an univariate function using the known problem of determining the circumcenter of a triangle. We also shown that the number of solutions of E3PNT is at most $6$.

@@

It is possible to make another reduction to obtain a degree-$6$ polynomial, this can lower the size of the companion matrix utilized on obtaining the roots of $h$ from $12^2=144$ to $6^2=36$.
As it has been mentioned before, an ellipse is symmetric with respect to its own major or minor axis. 
In other words, rotating an ellipse by $\theta \in [0, \pi)$ is the same as rotating the same ellipse by $\pi + \theta$.
On the other hand, on \autoref{chapter:definitions}, it was stated that the angle of the complex numbers $z$ and $-z$ has the same symmetry with each other as the aforementioned symmetry of ellipses:
\begin{equation*}
angle(-z) = \pi + angle(z).
\end{equation*}
From that, as $g(e^{i\theta})=\xi(\theta)$ for every $\theta \in [0, 2\pi)$, we conclude that $g(-z)=g(z)$. This implies that $h$ is, in fact, an even polynomial, or that $h(-z) = h(z)$ is true for every $z\in\mathbb{C}$:
\begin{align}
h(-z) = (-z)^6g(-z) = z^6g(z).
\end{align}
Therefore, all the odd degree coefficients of $h$ are $0$ and we can define the $6$-degree polynomial $f : \mathbb{C} \mapsto \mathbb{C}$ with the substitution $y=z^2$:
\begin{equation}
f(y) = \sum_{k=0}^{6} c_{2k} y^k.
\end{equation}
Then from every root $\hat{y}$ of $f$, two roots of $h$ can be obtained: $\sqrt{\hat{y}}$ and $-\sqrt{\hat{y}}$. As the angle of $-\sqrt{\hat{y}}$ is not between $[0, pi)$ we can ignore it. Note that the the square root of $\hat{y}$ does not need to be calculated, as only the angles are needed and they can be obtained by:
\begin{equation}
angle(\sqrt{z}) = angle(z)/2.
\end{equation}
Finally, using the QR algorithm mentioned on \autoref{chapter:definitions} a $\bigO(n^3)$ algorithm can be constructed for E3P. Because $n$ will always be $6$, we can say that we have a $\bigO(1)$ algorithm.

### 2019-12-18 defesa/tex/chapter/introducao.tex
b1

Two main types of optimal covering problems can be found in the literature: the Minimum Cover Problem, also known as just Set Cover Problem, and the Maximal Covering Problem \cite{karatas}.

@@

The second type of covering problems arose from the fact that covering almost all the demand set can be a lot cheaper than having to cover it all \cite{garcia}. This second type is known as \sigla{\MCLP}{Maximum Covering Location Problem}\footnote{Originally, the problem was defined using the word \textit{Maximal} on its title. In some sources, however, the word \textit{Maximum} is adopted.} and was introduced in \citeonline{church:1974}. In this first study, it is defined on a network with demand nodes, a facility set is also given and a solution maximizes the demand coverage satisfying the constraint that only a subset of the facilities is used. Just like the Minimum Cover, MCLP is a NP-Hard problem \cite{hatta:2013} and both deterministic, using integer programming in \citeonline{church:1974}, and heuristic methods in \citeonline{revelle:2008} have been proposed to solve it. A very complete survey of developments on this area can be found in \citeonline{mclp_review}.

In \citeonline{church:1984} a new kind of MCLP named \sigla{\PMCLP}{Planar Maximum Covering Location Problem} was introduced. Unlike its predecessor, this version of the problem was not defined on a network. Instead, on PMCLP the demand set and the facilities are located in $\R^2$ having a facility's coverage area be defined by a distance function. 
Initially, the Euclidean distance was considered and the idea behind the method proposed in \citeonline{church:1984} was to convert PMCLP into an instance of MCLP, and then utilize any of the previous developed exact methods to obtain a solution for PMCLP. This reduction was done by identifying a \sigla{CLS}{Candidate Locations Set} which represented the possible locations that needed to be evaluated for every facility, so the optimal solution could be found. From the CLS, a graph was built on which MCLP could be applied. Generating the CLS specifically for the case of Euclidean distance was done by the procedure called \sigla{CIPS}{Circle Intersect Points Set} which is described here on \autoref{chapter:pmclp}.

Also, some variations of PMCLP can be found in the literature: in \citeonline{younies} PMCLP was studied under the block norm distance; in \citeonline{mumtaz} a mean-shift algorithm for large scale\footnote{numerical experiments for up to $3000$ points} PMCLP was proposed; and in \citeonline{bansal} a version with partial coverage and demand and facility zones represented by rectangles was introduced.

The Euclidean norm PMCLP, which has a lot of results that can be applied for the elliptical PMCLP, is also found in the literature as the problem of maximization of points covered by a fixed number of unit disks \cite{cabello:2006}. 
Early works only tackled the one-disk version of the problem, in \citeonline{chazelle:1986} a $\bigO(n^2)$ algorithm, which still stands as the best in terms of run-time complexity, was proposed beating the prior $\bigO(n^2\log{n})$ algorithm created by \citeonline{drezner}.
The $m$ unit disks maximum covering was studied in \citeonline{cabello:2006} which had as its most important result a $(1-\epsilon)$-approximation algorithm which runs in $\bigO(n\log{n})$. To achieve its main goal, however, they developed a deterministic $\bigO(n^{2m-1}\log{n})$ algorithm which gets employed into their approximation scheme.
Additionally, in \citeonline{aronov:2008} one-disk maximum covering is proven to be 3SUM-HARD. This means that maximizing the amount of points covered by a disk is as hard as finding three real numbers that sum to zero among $n$ given real numbers.

Planar maximum covering with ellipses differs from its disks counterpart only in the shape of the facility's coverage area. The main motivation to study this modified version is that cellphone towers can have elliptical shaped coverage area, so in order to determine what are the best locations to place $m$ cellphone towers to maximize the amount of the population covered by its signal, an elliptical PMCLP is better suited \cite{canbolat}.

Only two articles have been found published in the literature that study this problem. In \citeonline{canbolat}, a mixed-integer non-linear programming method was proposed as a first approach to the problem. For some instances the method took too long and did not find an optimal solution. For this reason a heuristic method was developed using a technique called Simulated Annealing. Solutions for the instances that timed-out with the first method were then obtained. The problem was further explored in \citeonline{andreta} which proposed a deterministic method that showed better performance obtaining optimal solutions for the instances which the first method could not. Also, in \citeonline{andreta}, a version of the problem where every ellipse can be freely rotated was introduced and an exact method, which could not find optimal solutions for large instances, and a heuristic method were proposed for it. Despite the similarities, none of the works cited above base their development on the maximum covering with disks algorithms found in the literature.

This work is structured in the following way: \autoref{chapter:definitions} introduces some definitions and results that are used throughout the next chapters; in \autoref{chapter:pmclp}, the maximum covering by disks problem is studied and a $\bigO(n^{2m})$ algorithm is proposed; in \autoref{chapter:ellipses}, the maximum covering by ellipses is introduced and the algorithm for the disks case is adapted for it; finally, \autoref{chapter:future_work} presents what is left as future work. Also, \autoref{chapter:ellipses_intersection} determines with detail the intersection of two ellipses, which is used in the algorithm developed in \autoref{chapter:ellipses}.

>>>>

aTwo main types of optimal covering problems can be found in the literature: the Minimum Cover Problem, also known as just Set Cover Problem, and the Maximal Covering Problem \cite{karatas}.

@@

The second type of covering problems arose from the fact that covering almost all the demand set can be a lot cheaper than having to cover it all \cite{garcia}. This second type is known as \sigla{\MCLP}{Maximal Covering Location Problem} and was introduced in \cite{church:1974}. In this first study, it is defined on a network with demand nodes, a facility set is also given and a solution maximizes the demand coverage satisfying the constraint that only a subset of the facilities is used. Just like the Minimum Cover, MCLP is a NP-Hard problem \cite{hatta:2013} and both deterministic, using integer programming \cite{church:1974}, and heuristic methods \cite{revelle:2008} have been proposed to solve it.

In \cite{church:1984} a new kind of MCLP named \sigla{\PMCLP}{Planar Maximal Covering Location Problem} was introduced. This version of the problem was not defined on a network, instead the demand set and the facilities are located in $\R^2$ having the coverage area of a facility be defined by a distance function. PMCLP is said to have been studied under Euclidean and rectilinear distance functions \cite{younies}. The Euclidean norm PMCLP, which has a lot of results that can be applied for the elliptical PMCLP, is also found in the literature as the problem of maximization of points covered by a fixed number of unit disks \cite{cabello:2006}. 
Early works only tackled the one-disk version of the problem, in \cite{chazelle:1986} a $\bigO(n^2)$ algorithm, which still stands as the best in terms of run-time complexity, was proposed beating the prior $\bigO(n^2\log{n})$ algorithm created by \cite{drezner}.
The $m$ unit disks maximal covering was studied in \cite{cabello:2006} which had as its most important result a $(1-\epsilon)$-approximation algorithm which runs in $\bigO(n\log{n})$. To achieve its main goal, however, they developed a deterministic $\bigO(n^{2m-1}\log{n})$ algorithm which gets employed into their approximation scheme.
Additionally, in \cite{aronov:2008} one-disk maximal covering is proven to be 3SUM-HARD. This means that maximizing the amount of points covered by a disk is as hard as finding three real numbers that sum to zero among $n$ given real numbers.

Planar maximal covering with ellipses differs from its disks counterpart only in the shape of the facility's coverage area. The main motivation to study this modified version is that cellphone towers can have elliptical shaped coverage area, so in order to determine what are the best locations to place $m$ cellphone towers to maximize the amount of the population covered by its signal, an elliptical PMCLP is better suited \cite{canbolat}. Only two articles have been found published in the literature that study this problem. In \cite{canbolat}, a mixed-integer non-linear programming method was proposed as a first approach to the problem. For some instances the method took too long and did not find an optimal solution. For this reason a heuristic method was developed using a technique called Simulated Annealing. Solutions for the instances that timed-out with the first method were then obtained. The problem was further explored in \cite{andreta} which proposed a deterministic method that showed better performance obtaining optimal solutions for the instances which the first method could not. Also, in \cite{andreta}, a version of the problem where every ellipse can be freely rotated was introduced and an exact method, which could not find optimal solutions for large instances, and a heuristic method were proposed for it. Despite the similarities, none of the works cited above base their development on the maximal covering with disks algorithms found in the literature.

This work is structured in the following way: \autoref{chapter:definitions} introduces some definitions and results that are used throughout the next chapters; in \autoref{chapter:pmclp}, the maximal covering by disks problem is studied and a $\bigO(n^{2m})$ algorithm is proposed; in \autoref{chapter:ellipses}, the maximal covering by ellipses is introduced and the algorithm for the disks case is adapted for it; finally, \autoref{chapter:future_work} presents what is left as future work. Also, \autoref{chapter:ellipses_intersection} determines with detail the intersection of two ellipses, which is used in the algorithm developed in \autoref{chapter:ellipses}.
